Reproducible Research: Course Project 1
=============================================

## First, let's see what is our directory, set our directory, etc.
```{r}
getwd()
setwd("/Users/Simon/Coursera_R_programming/Course project 1 - Reproducible research")
```

## Second, let's load the libraries that we will use
```{r, results = 'hide'}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Third, let's load the dataset to be analysed
### 1. Load the dataset using read.csv()
```{r}
if(!file.exists("activity.csv")) {
        unzip("activity.csv")
}
activity.data <- read.csv("activity.csv")
```

### 2. Before any data analysis, let's see how the data look like
```{r}
head(activity.data)
summary(activity.data)
str(activity.data)
```

## We begin with the data analysis
```{r}
steps.per.day <- aggregate(steps ~ date, activity.data, sum)
```

## What is the mean total number of steps taken per day?
### 1. Create histogram of total number of steps taken per day
```{r}
hist(steps.per.day$steps, main = "Total steps per day", col = "royalblue", xlab = "Number of steps")
```

### 2. Compute mean and median of steps taken per day
```{r}
mean.steps <- mean(steps.per.day$steps)
median.steps <- median(steps.per.day$steps)

mean.steps
median.steps
```

## What is the average daily activity pattern?
### 1. Calculate average steps for each interval for all days
```{r}
steps.per.interval <- aggregate(steps ~ interval, activity.data, mean)
```

### 2. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
plot(steps.per.interval$interval, steps.per.interval$steps, type="l", xlab = "Interval", ylab = "Number of Steps",main = "Average Number of Steps per Day by Interval")
```

### 3. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max.interval <- steps.per.interval[which.max(steps.per.interval$steps), 1]
```

## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
total.na <- sum(!complete.cases(activity.data))
```

### 2. Devise a strategy for filling in all of the missing values in the dataaset
### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
activity.data.imputed <- activity.data
activity.data.imputed$steps <- impute(activity.data$steps, fun = mean)
```

### 4. Create a histogram of the total number of steps taken each day (imputed dataset)
```{r}
steps.per.day.imputed <- aggregate(steps ~ date, activity.data.imputed, sum)
hist(steps.per.day.imputed$steps, main = "Total steps per day (imputed)", col = "red", xlab = "Number of steps")
```

### 5. Compute new mean and median for imputed dataset
```{r}
mean.steps.imputed <- mean(steps.per.day.imputed$steps)
median.steps.imputed <- median(steps.per.day.imputed$steps)

mean.steps.imputed
median.steps.imputed
```

### 6. Calculate difference between original and imputed datasets
```{r}
mean.diff <- mean.steps.imputed - mean.steps
median.diff <- median.steps.imputed - median.steps

mean.diff
median.diff
```

### 7. Calculate total difference
```{r}
total.diff <- sum(steps.per.day.imputed$steps) - sum(steps.per.day$steps)

total.diff
```

## Are there differences in acticuty patterns between weekdays and weekends?
### 1. Create a new factor variable in the dataset with two levels, "weekday" and "weekend", indicating whether a given date is a weekday or weekend day.
```{r}
activity.data.imputed$date.type <- ifelse(as.POSIXlt(activity.data.imputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
averaged.activity.data.imputed <- aggregate(steps ~ interval + date.type, data = activity.data.imputed, mean)

ggplot(averaged.activity.data.imputed, aes(interval, steps)) +
        geom_line() + 
        facet_grid(date.type ~ .) + 
        xlab("5-minute interval") + 
        ylab("avarage number of steps")
```

